Title: Fish do not need Visas: the transboundary nature of world fished species (70/96 character maximum)

### 2019-11-11 Manuscript_Draft.Rmd
Updates to intro

---
title: "Title: Fish do not need Visas: the transboundary nature of world fished species (70/96 character maximum)"
authors: "Authors: Juliano Palacios-Abrantes^1^, Gabriel Reygondeau^1,2^, William W.L. Cheung^1^"
date: "Affiliations: ^1^ Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada, ^2^ Yale"
# subtitle: "Fish do not need Visas (16/40 character maximum)"
output:
  pdf_document:
    fig_caption: yes
  word_document:
    reference_docx: ./Manuscript_Instructions/science_template.docx
geometry: margin=1in
editor_options:
  chunk_output_type: console
bibliography: ./References/FishForVisa_Ref.bib
csl: ./References/science.csl
---

@@

<!-- # Introduction (~600 words) -->

Species are not randomly distributed in the world. The area in space and time where a stable population lives and reproduces is determined by the species' fitness response to a series of biotic and abiotic factors [@Hunch] and their influence in the evolutionary process [@REF]. Species distributions are key elements in marine resources management as human-made spatial boundaries are a major part of the management system [@Song:2017iua]. However, incorporating the concept of species distribution has not allays been the case in fisheries management, nor policy [@Alexa]. The delimitation of Economic Exclusive Zones (EEZs) in the arly 80's virtually splitted the distribution of many species, creating shared species between nations, and adding another level of complexity to fisheries management. However, 40 years after their establishment, the question of how many shared species exists in the world based on their distribution is still unanswered. In here, we take the most detailed database of marine fished species to show that many species are shared, representing an important contribution to national economies and food security, and highlighting the importance to acknowledge this in the policy arena.

@@

There is a gap in assessing transboundary species in the world based on their spatial distribution. It was estimated in 1997 that there could be up to 1,500 transboundary fish species in the world, however, such estimation lacked a proper assessment due to limited information at the time [@Caddy:1997ue]. More recently, it was estimated there were 344 shared species, representing a capture of 34.2 × 10^6^ t and a global landed value of USD 30.7 × 10^9^ [@Teh:2015gd]. While this represents a good baseline to start, the analysis was based on a literature review of shared species and did not include a mechanistic way of determine the transboundary nature of the species [@Teh:2015gd]. Therefore, 40 years after UNCLOS, the question of how many transboundary species exists in the world is still unanswered. This is of uppermost need as the fisheries management approach differs substantially between EEZ-discreet and shared species [@Mills].

 <!-- **Results and discussion** (1000 words) -->

We combined a series of datasets and species distribution models to estimate the number of transboundary fished species and their contribution to food security and economics of the world. We adopted the UN definition of transboundary stocks (shared by neighboring EEZs) but carried the analysis at the species level, rather than stocks (populations within a species), due to lack of stock-specific spatial and ecological information on all of the fished species. Our results suggest that there are `r Total_trans_n` transboundary species in the world, wich is **something** related to what Caddy (1997) estimated and more than double of that of Teh and Sumaila (2010) [@Teh].

Tranboundary fisheries are quite important worldwide in terms of both landings and revenue as they are responsable for XXX tones of landings generating USD `r Total_Trans_Rev` million in fishing revenue. This represents XX% of the catch and `r round(Per_Trans_Rev,2)`% of total revenue generated by those fisheries in all EEZs of the world. All these values are higher than previouslly reported by Teh and Sumaila (2010) up to 20006 where the maximum amount was 45% in the mid-1990s [@Teh]. The importance of these species exceeds in EEZs like the United States, Peru and Norway where fewer species are responsable for higer landing values (Fig 1).

@@

Transboundary fisheries are vast in regions whith historic fisheries-related conflicts [@Spikjers]. Regional variation show higher frequency of transboundary species within EEZs along the South China sea, western central Atlantic, and the Mediterranean sea, followed by the Eastern north Atlantic (Fig 1). Some of these regions have been identified as most conflictive since 1970s with conflicts arrousing to even armed conflicts. Moreover, as climate change shifts species distributions and new transboundary species arrose [@Pinsky] as seen in past cases like the Pacific salmon [@REF], or the Macrrell waars [@Spijkers]. The joint management of shared stocks could be key to not only sustain fisheries into the future but also avoid international conflict [@REF].

Management of shared stocks have to be something. Although catch trends do not reflect stock status [@REf_Haters], they can inform provide a panoram of the situation when no other data is available [@Pauly]. In the case of transboundary species, we did not find a strong pattern suggesting a specific case (Fig 2).

>>>>

---
title: "Title: Fish do not need Visas: the transboundary nature of world fished species (70/96 character maximum)"
authors: "Authors: Juliano Palacios-Abrantes^1^, Gabriel Reygondeau^1,2^, William W.L. Cheung^1^"
date: "Affiliations: ^1^ Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada, ^2^ Yale"
# subtitle: "Fish do not need Visas (16/40 character maximum)"
output:
  word_document:
    reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: console
bibliography: ./References/FishForVisa_Ref.bib
csl: ./References/science.csl
---

@@

<!-- # Introduction (~555/500 words) -->

Species are not randomly distributed in the world. The area in space and time where a stable population lives and reproduces is determined by the species' fitness response to a series of biotic and abiotic factors [@Hunch] and their influence in the evolutionary process [@REF]. Species distributions are key elements in marine resources management as human-made spatial boundaries are a major part of the management system [@Song:2017iua]. However, incorporating the concept of species distribution has not allays been the case in fisheries management, nor policy [@Alexa]. The delimitation of Economic Exclusive Zones (EEZs) in the arly 80's virtually splitted the distribution of many species, creating shared species between nations, and adding another level of complexity to fisheries management. However, 40 years after their establishment, the question of how many shared species exists in the world based on their distribution is still unanswered.

@@

There is a gap in assessing transboundary species in the world based on their spatial distribution. It was estimated in 1997 that there could be up to 1,500 transboundary fish species in the world, however, such estimation lacked a proper assessment due to limited information at the time [@Caddy:1997ue]. More recently, it was estimated there were 344 shared species, representing a capture of 34.2 × 10^6^ t and a global landed value of USD 30.7 × 10^9^ [@Teh:2015gd]. While this represents a good baseline to start, the analysis was based on a literature review of shared species and did not include a mechanistic way of determine the transboundary nature of the species [@Teh:2015gd]. And so, despite recent research higlighting the interconnection of marine species [@Nandini], transboundary species are yet to be identified at a global scale. This is of uppermost need as the fisheries management approach differs substantially between EEZ-discreet and shared species [@Mills].

 <!-- **Methods summary** (~140/250 words) -->

In here, we overlap the known distribution of **968** marine fish species with XXX Economic Exclusive Zones of  the wrold to map the global number of transboundary species and and their contribution to gloabal catch and landed value. We adopted the UN definition of transboundary stocks (shared by neighboring EEZs) but carried the analysis at the species level, rather than stocks (populations within a species), due to lack of stock-specific spatial and ecological information on all of the fished species [@TEH].

We combined multiple species distribution data-sources and developed a series of criterias to determine if a species was transboundary. Current species distributions were obtained from an observational database of world marine species [@REFNEreus], an environmental niche model based on environmental variables and multiple algorithms [@REF], and a species distribution model based on species traits [@Zeler]. The first criteria to determine if a species is transboundary was to Criteria #1. Followed by

our study provides a first empirical approximation to understand the true transboundary nature of the world fisheries, a world that is deeply coonected by larval dynamics [@nandini], fishing fleets [@Kurosawa_Juanito], and shared management [@RFMOS].

<!-- **Results and discussion** (~500/1000 words) -->

As Caddy (1997) pointed out, virtaully all coastal nations share at least one transboundary species with their neighbours, however, some countries share over 100 species with sourronding EZS (Fig. 1). According ot our model, there are a total of `r Total_trans_n` transboundary species in the world, more than double of what had been previouslly estimated [@Teh:2015gd]. Transboundary fisheries are vast in regions whith historic fisheries-related conflicts [@Spikjers]. Regional variation show higher frequency of transboundary species within EEZs along the South China sea, western central Atlantic, and the Mediterranean sea, followed by the Eastern north Atlantic (Fig 1). Some of these regions have been identified as most conflictive since 1970s with conflicts arrousing to even armed conflicts. Conflicts over shared resources in these regions include... EXAMPLES.

Conflicts in these regions will be exacerabted by climate change. As species shift their known distributions to cope with a changing ocean, new transboundary species will arrose, challenging the management of shared stocks worldwide [@Pinsky]. Climate change has previouslly caused conlfict between important fishing nations such as Canada and the United States over Pacific salmon (*Oriche spp.*) back in the 1980's [@REF], France and Italy over XXX in the 1990's, and more recnetly Iralnd and the European Union over Macrrell (*species*) [@Spijkers]. The identification of transboundary species is a first step to determine the transboundary nature of a shared stock and establish a joint management plan, key to not only sustain fisheries into the future but also avoid international conflict [@REF].

Tranboundary fisheries are quite important worldwide in terms of both landings and revenue. In the last ten years, the global feleet has landed an average of XXX tones, generating USD `r Total_Trans_Rev` million in fishing revenue. This represents XX% of the gloabl catch and `r round(Per_Trans_Rev,2)`% of total revenue generated by those fisheries in all EEZs of the world. All these values are higher than previouslly reported by Teh and Sumaila (2010) up to 20006 where the maximum amount was 45% in the mid-1990s [@Teh]. The importance of these species exceeds in EEZs like the United States, Peru and Norway where fewer species are responsable for higer landing values (Fig 1).

@@

Management of shared stocks have to be improved if we are to achieve the sustainable goals of SDGs... Although catch trends do not reflect stock status [@REf_Haters], they can inform provide valuable information of the situation when no other data is available [@Pauly]. Based on our catch trend clasification, XX% of countries are harvesting their shared stocks between 10 and 50% of the maximum historical catch (Cat, Max.Eplo), followed by XX% harvesting between 10-50% of the maximum historic catch (Cat.), and XX% harvesting something else (Cat.)  (Fig 2). Several emipiracal [@REF;@REF;@REF] and modelled studies [@REF;@REF;@REF;] have shown that cooperation will have beeter outcomes for shared fisheries management. Blablablabala
